---
title: "WaterUseBySector"
author: "Team NatCapAccounting"
date: "7/21/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Water Use by Sector
This project pulls data from the USGS water use data server for the years 2000, 2005, and 2010, then combines these data into a single table from which a user can view the water use by sector for a selected state. 

### Load libraries
```{r echo=FALSE}
#load libraries
library(tidyr)
library(dplyr)
library(ggplot2)
```

### Read in the USGS data
```{r}
#read in usgs use data
df2000 <- read.delim('https://water.usgs.gov/watuse/data/2000/usco2000.txt',na='')
df2005 <- read.delim('https://water.usgs.gov/watuse/data/2005/usco2005.txt',na='')
df2010 <- read.delim('https://water.usgs.gov/watuse/data/2010/usco2010.txt',na='')
head(df2010)
```

### Rename columns to more readable names 
Year 2000...
```{r}
#Rename rename freshwater total columns
df2000 <- df2000 %>%
  rename("Population" = "TP.TotPop") %>%
  rename("Public" = "PS.WFrTo") %>%
  rename("Domestic" = "DO.WFrTo") %>%
  rename("Industrial" = "IN.WFrTo") %>%
  rename("Irrigation" = "IT.WFrTo") %>%
  rename("Aquaculture" = "LA.WFrTo") %>%
  rename("Livestock" = "LS.WFrTo") %>%
  rename("Mining" = "MI.WFrTo") %>%
  rename("Thermoelectric" = "PT.WFrTo") %>%
  rename("Total" = "TO.WFrTo") %>%
#Remove all other supply columns
  select(-contains("."))
#Add year column equal to 2000
df2000['YEAR'] = 2000
```

Year 2005...
```{r}
#Rename rename freshwater total columns
df2005 <- df2005 %>%
  rename("Population" = "TP.TotPop") %>%
  rename("Public" = "PS.WFrTo") %>%
  rename("Domestic" = "DO.WFrTo") %>%
  rename("Industrial" = "IN.WFrTo") %>%
  rename("Irrigation" = "IR.WFrTo") %>%
  rename("Aquaculture" = "LA.WFrTo") %>%
  rename("Livestock" = "LS.WFrTo") %>%
  rename("Mining" = "MI.WFrTo") %>%
  rename("Thermoelectric" = "PT.WFrTo") %>%
  rename("Total" = "TO.WFrTo") %>%
  #Remove all other supply columns
  select(-contains(".")) %>%
  select(-X)
df2005['YEAR'] = 2005
```

Year 2010...
```{r}
#Rename rename freshwater total columns
df2010 <- df2010 %>%
  rename("Population" = "TP.TotPop") %>%
  rename("Public" = "PS.WFrTo") %>%
  rename("Domestic" = "DO.WFrTo") %>%
  rename("Industrial" = "IN.WFrTo") %>%
  rename("Irrigation" = "IR.WFrTo") %>%
  rename("Aquaculture" = "AQ.WFrTo") %>%
  rename("Livestock" = "LI.WFrTo") %>%
  rename("Mining" = "MI.WFrTo") %>%
  rename("Thermoelectric" = "PT.WFrTo") %>%
  rename("Total" = "TO.WFrTo") %>%
  #Remove all other supply columns
  select(-contains(".")) %>%
  select(-COUNTY) %>%
  select(-X)
head(df2010)
```


### Append all yearly data into a single table & clean up
```{r}
#Combine all tables
dfUse = bind_rows(df2000, df2005, df2010)

#Remove old tables
rm(df2000, df2005, df2010)

head(dfUse)
```


### Summarize data by state
```{r}
#Generate the table of year x summed usage & supply
byStates <- dfUse %>% 
  #Group by time and state
  group_by(YEAR,STATE) %>%
  #Summarize all records
  summarize("Population" = sum(Population),
            "Public" = sum(Public), 
            "Domestic" = sum(Domestic),
            "Industrial" = sum(Industrial),
            "Irrigation" = sum(Irrigation),
            "Aquaculture" = sum(Aquaculture),
            "Livestock" = sum(Livestock),
            "Mining" = sum(Mining),
            "Thermoelectric" = sum(Thermoelectric),
            "Total" = sum(Total))
head(byStates)
```

### Gather data into a tidier format
```{r}
#Gather the data
stateGather <-  gather(byStates, key=Sector, value=Withdrawal, Public:Thermoelectric)
head(stateGather)
```

### Filter for a specific state
```{r}
dfState = stateGather %>%
  filter(STATE == 'MD')
  
head(dfState)
```

### Plot data for a selected state
```{r}

thePlot <- ggplot(data = dfState,
       aes(x=as.numeric(YEAR),y=log10(Withdrawal),color=Sector)) +
  geom_line() +
  geom_point() +
  labs(title="State withdrawal by sector", 
       x="YEAR", 
       y="Withdrawal (MGal/year)") +
  scale_x_continuous(limits = c(2000,2010),
                     breaks = c(2000,2005,2010))
         
thePlot <- thePlot + theme(plot.title = element_text(hjust = 0.5))
thePlot
```

