---
title: "WaterUseBySector"
author: "Team NatCapAccounting"
date: "7/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Water Use by Sector
This project pulls data from the USGS water use data server for the years 2000, 2005, and 2010, then combines these data into a single table from which a user can view the water use by sector for a selected state. 

### Load libraries
```{r echo=FALSE}
#load libraries
library(tidyr)
library(dplyr)
library(ggplot2)
```

### Read in the USGS data
```{r}
#read in usgs use data
df2000 <- read.delim('https://water.usgs.gov/watuse/data/2000/usco2000.txt')
df2005 <- read.delim('https://water.usgs.gov/watuse/data/2005/usco2005.txt')
df2010 <- read.delim('https://water.usgs.gov/watuse/data/2010/usco2010.txt')
```

### Rename columns to more readable names 
```{r}
#Rename rename freshwater total columns
df00 <- df2000 %>%
  rename("Population" = "TP.TotPop") %>%
  rename("Public" = "PS.WFrTo") %>%
  rename("Domestic" = "DO.WFrTo") %>%
  rename("Industrial" = "IN.WFrTo") %>%
  rename("Irrigation" = "IT.WFrTo") %>%
  rename("Aquaculture" = "LA.WFrTo") %>%
  rename("Livestock" = "LS.WFrTo") %>%
  rename("Mining" = "MI.WFrTo") %>%
  rename("Thermoelectric" = "PT.WFrTo") %>%
  rename("Total" = "TO.WFrTo") %>%
#Remove all other supply columns
  select(-contains("."))
#Add year column equal to 2000
df00['YEAR'] = 2000
```


### Append all yearly data into a single table
```{r}

```

### Summarize data by state
```{r}

```

### Gather data into a tidier format
```{r}

```

### Plot data for a selected state
```{r}

```

